Hydras \& Co.

### 2022-05-24 doc/ks-chapter.tex
small corrections

Our proofs are  constructive and require no axioms: they are  closed terms of the CIC, and are mainly composed on function definitions and proofs of properties of these functions. 
They  borrow much theoretical material from Kirby and Paris, although they do not use any knowledge about Peano arithmetic nor model  theory.  The combinatorial arguments we use and implement
come from 
 an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
 by L.~Kirby and J.~Paris. % on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
 Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
rounds of hydra battles. 
Nevertheless, they have the same global structure as the simple proofs described in
sections~\vref{omega-case} and \vref{omega2-case}. 
We invite the reader to compare the three proofs step by step, lemma by lemma.

@@

@@

First, let us prove the following lemma (numbered \texttt{2.4} in Ketonen and Solovay's article).

We prove this lemma by transfinite induction on $\alpha$.
Let us consider a path $\alpha \xrightarrow [i]{} \beta$ $(i>0)$. Its first step is
the pair $(\alpha,\canonseq{\alpha}{i})$, We have $\canonseq{\alpha}{i}<\alpha$ and
$\canonseq{\alpha}{i} \xrightarrow [i]{} \beta$. 
Let $n$ be any natural number such that $n>i$.
By the induction hypothesis, there exists a path $\canonseq{\alpha}{n} \xrightarrow[i]{} \beta$.
\begin{itemize}
\item  If $\alpha$ is a successor ordinal $\gamma+1$, then $\canonseq{\alpha}{n} =
\canonseq{\alpha}{i}=\gamma$. Thus we have a path 
$\alpha  \xrightarrow [n]{}  \gamma \xrightarrow [n]{} \beta$
\item If $\alpha$ is a limit ordinal, we apply Theorem \texttt{KS\_thm\_2\_4} (see above).

@@

Why is \texttt{Cor12} so useful? 
Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying corollary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
where $n$ is the maximum of the indices $i$ met in the induction.

@@

Since comparison on \texttt{T1} is decidable, one can compute the last step $\gamma$ of the standard path from $(\alpha,n+1)$  such that $\beta\leq \gamma$.
Let $l$ be the length of the path from $\alpha$ to $\gamma$.  
This step of the proof is illustrated in figure~\vref{fig:belle-preuve-2}.

>>>>

Our proofs are  constructive and require no axioms: they are  closed terms of the CIC, and are mainly composed on function definitions and proofs of properties of these functions. 
They  borrow much theoretical material from Kirby and Paris, although they do not use any knowledge about Peano arithmetic nor about model  theory.  The combinatorial arguments we use and implement
come from 
 an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
 by L.~Kirby and J.~Paris. % on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
 Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
rounds of hydra battles. 
Nevertheless, they have the same global structure as the simple proofs described in
sections~\vref{omega-case} and \vref{omega2-case}. 
We invite the reader to compare the three proofs step by step, lemma by lemma.

@@

\texttt{Cor12} is a consequence of the following theorem (numbered \texttt{2.4} in Ketonen and Solovay's article), proven by transfinite induction on $\alpha$.

@@

\texttt{Cor12} is also proved by transfinite induction on $\alpha$.  Let us give  a sketch of its proof \footnote{This proof sketch is a slight simplification of the formal proof script:
  The strictly positive indexes $i$ and $n$ stand for the
  terms \texttt{(S i)} and \texttt{(S n)}. We do not explicit  the (simpler) case where the considered path is made of only one step.}

Let us consider a path $\alpha \xrightarrow [i]{} \beta$ $(i>0)$. Its first step is
the pair $(\alpha,\canonseq{\alpha}{i})$, We have $\canonseq{\alpha}{i}<\alpha$ and
$\canonseq{\alpha}{i} \xrightarrow [i]{} \beta$. 
Let $n$ be any natural number such that $n>i$.
By the induction hypothesis, there exists a path $\canonseq{\alpha}{n} \xrightarrow[i]{} \beta$.
\begin{itemize}
\item  If $\alpha$ is a successor ordinal $\gamma+1$, then $\canonseq{\alpha}{n} =
\canonseq{\alpha}{i}=\gamma$. Thus we have a path 
$\alpha  \xrightarrow [n]{}  \gamma \xrightarrow [n]{} \beta$
\item If $\alpha$ is a limit ordinal, we apply Theorem \texttt{KS\_thm\_2\_4} (see above).

@@

Why is \texttt{Cor12} so useful? 
Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some natural number. Applying corollary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
where $n$ is the maximum of the indices $i$ met in the induction.

@@

Since comparison on \texttt{T1} is decidable, one can compute the last step $\gamma$ of the standard path from $(\alpha,n+1)$  such that $\beta\leq \gamma$.
Let $l$ be the length of the path from $\alpha$ to $\gamma$.  
This step of the proof is illustrated in figure~\vref{fig:belle-preuve-2} \footnote{Please note that in these figures, smaller ordinals are represented on the right side!}.
